---
title: "Exercise4Q1(Jonathan)"
author: "Jyun-Yu Cheng ,YiJi Gao"
date: "5/1/2022"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Question 1 : Clustering and PCA
```{r message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(readr)
library(reshape2)
library(modelr)
library(RCurl)
library(mosaic)
library(LICORS) # For K-Means++  
library(ggplot2)
library(ggpubr)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Read the data
wine_original = read.csv(
  "https://raw.githubusercontent.com/jgscott/ECO395M/master/data/wine.csv"
  ,stringsAsFactors=T)
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
# BC of the color variable is not a numeric variable
wine_data = wine_original %>% 
  select(-quality,-color) %>% 
  scale(center=TRUE, scale=TRUE) 
# transfer the wine_data to data frame
wine_data = as.data.frame(wine_data)
# view(wine_data)
```

## Run the PCA and Clustering
## Clustering part
```{r message=FALSE, warning=FALSE, echo=FALSE}
# Clustering: K-Means++ #
## Use K-Means++ to divide data points' color into  2 categories 
wine_clustering_color = kmeanspp(wine_data, k=2, nstart=20)

## Use K-Means++ to divide data points' quality into  7 categories
wine_clustering_quality = kmeanspp(wine_data, k= 7, nstart=20)

wine_original$k_color_cluster = wine_clustering_color$cluster
wine_original$k_quality_cluster = wine_clustering_quality$cluster
# Build a new data frame
wine_new = data.frame(wine_data, color = wine_original$color, 
                          quality = wine_original$quality,
                          clust_color = wine_clustering_color$cluster,
                          clust_quality = wine_clustering_quality$cluster)
# We had used clustering to divide the data points successfully 
```



## Assess : Clustering  
### Color
```{r message=FALSE, warning=FALSE, echo=FALSE}
# Assess clustering results
 
# We use figure to assess the clustering results
# Combine the wine_new 12 to 15 columns and 
# Used stack to transform data available as separate columns in a data frame 
wine_data_clustering = cbind(wine_new[12:15], stack(wine_new[1:11]))
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
ggplot(wine_data_clustering) +
  geom_boxplot(aes(x=color, y=values)) +
  facet_wrap(~ind, ncol=3) +
  labs(y = "Chemical Properties' Values",
       x="Real Color")
```
According to this figure, we can see that we may distinguish wines by fixed.acidity and volatile.acidity


```{r message=FALSE, warning=FALSE, echo=FALSE}
### Labels emerged naturally from clustering

ggplot(wine_new) + 
  geom_point(aes(x=volatile.acidity, y=total.sulfur.dioxide,
                 col=factor(clust_color))) +
  labs(color='Generated Cluster') 
```
By clustering, We can see that one group is on the left, and the other group is on the right.

```{r message=FALSE, warning=FALSE, echo=FALSE}
ggplot(wine_new) + 
  geom_point(aes(x=volatile.acidity, y=total.sulfur.dioxide,
                 col=factor(color))) +
  labs(color='True Color')
```

This is the true color group of wines, compare to the above figure(clustering), we can see that clustering can help us distinguish the color of wine.




### Quality 
Now we are going to see that whether clustering works for  quality
Now we see the true quality figure
```{r message=FALSE, warning=FALSE, echo=FALSE}
### Data visualization
# this is true quality figure
ggplot(wine_data_clustering) +
  geom_boxplot(aes(x=factor(quality), y=values)) +
  facet_wrap(~ind, ncol=3) +
  labs(y = "Chemicals' values", x="True Quality")
```
We can see that these figures are very disordered, so we can't use it to determine
Because it is difficult to see which one can help us to distinguish, so we decide to use mean value of them to see which may work.
```{r message=FALSE, warning=FALSE, echo=FALSE}
wine_data_clustering %>%
  group_by(quality, ind) %>%
  summarise(median_feature = median(values)) %>%
  ggplot() +
  geom_point(aes(x=factor(quality), y=median_feature)) +
  facet_wrap(~ind, ncol=3) +
  labs(y = "Chemicals' Median Value",
       x="True Quality")
# may differed by alcohol and density, citric.acid and Free.sulfur.dioxide
```
We can see that there are 4 properties(alcohol,density,citric.acid and Free.sulfur.dioxide) may be used to distinguish the quality because they looks like they have some trends.


#### clustering labels 
```{r message=FALSE, warning=FALSE, echo=FALSE}
### Labels emerge naturally from clustering


# density & alcohol
clustering_1 = ggplot(wine_new) + 
  geom_point(aes(x=density, y=alcohol,
                 col=factor(clust_quality))) +
  labs(color="Generated Cluster")
# clustering_1
# vol and fsd
clustering_2 = ggplot(wine_new) + 
  geom_point(aes(x=volatile.acidity, y=free.sulfur.dioxide,
                 col=factor(clust_quality))) +
  labs(color="Generated Cluster")

# density and fsd
clustering_3 = ggplot(wine_new) + 
  geom_point(aes(x=density, y=free.sulfur.dioxide,
                 col=factor(clust_quality))) +
  labs(color="Generated Cluster")

#alcohol and fsd
clustering_4 = ggplot(wine_new) + 
  geom_point(aes(x=alcohol, y=free.sulfur.dioxide,
                 col=factor(clust_quality))) +
  labs(color="Generated Cluster")

# intergrate all clustering to one figure
ggarrange(clustering_1, clustering_2, clustering_3, clustering_4, 
          ncol = 2, nrow=2, common.legend = TRUE,
          legend = "right") 
```
We can see that we  still can't catch the difference among wines accurately,

#### Orginal label
Now we see the true quality of wines
```{r message=FALSE, warning=FALSE, echo=FALSE}

clust_1_original = ggplot(wine_new) + 
  geom_point(aes(x=density, y=alcohol,
                 col=factor(quality))) +
  labs(color="True Quality")

clust_2_original = ggplot(wine_new) + 
  geom_point(aes(x=volatile.acidity, y=free.sulfur.dioxide,
                 col=factor(quality)))  +
  labs(color="True Quality")

clust_3_original = ggplot(wine_new) + 
  geom_point(aes(x=density, y=free.sulfur.dioxide,
                 col=factor(quality)))  +
  labs(color="True Quality")

clust_4_original = ggplot(wine_new) + 
  geom_point(aes(x=alcohol, y=free.sulfur.dioxide,
                 col=factor(quality))) +
  labs(color="True Quality")

ggarrange(clust_1_original, clust_2_original, clust_3_original, clust_4_original, 
          ncol = 2, nrow=2, common.legend = TRUE,
          legend = "right") 
```
Compare the above 2 figures, we can know that it's very difficult to distinguish the quality of wines by use clustering.






## PCA part

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Use PCA to to divide data points' into categories
Wine_PCA = prcomp(wine_data, rank = 11)
PCA_Model = predict(Wine_PCA, wine_data)
wine_loadings = Wine_PCA$rotation %>%
  as.data.frame %>%
  rownames_to_column('features')
wine_scores = Wine_PCA$x %>%
  as.data.frame() %>%
  rownames_to_column('wine_code')

wine_new = wine_new %>% rownames_to_column('wine_code')
```


### colors 
```{r message=FALSE, warning=FALSE, echo=FALSE}
wine_new = merge(wine_new, wine_scores, by = 'wine_code') 
wine_data_pca = melt(wine_new, id.var = colnames(wine_new)[1:16],
                          variable.name = 'PC')
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
##找出所有的PC的影響
ggplot(wine_data_pca) +
  geom_boxplot(aes(x=color, y=value)) +
  facet_wrap(~PC) +
  labs(y = "Principal Components' Values",
       x="True Color")
```
From this figure, we can see that PC1 and PC2 may have some trends, so we may use it to distinguish color of wines

Then we plot it into one figure to see their abilities of distinguish

```{r message=FALSE, warning=FALSE, echo=FALSE}
ggplot(wine_new) +
  geom_point(aes(x=PC1, y=PC2, color=factor(color))) +
  labs(y = "PC2", x="PC1",
       color='True Color')
```
We can see that the points in the left are usually red wines, in the right are usually white wines. So PCA can distinguish the color of wines successfully. 


### quality 
```{r message=FALSE, warning=FALSE, echo=FALSE}
ggplot(wine_data_pca) +
  geom_boxplot(aes(x=factor(quality), y=value)) +
  facet_wrap(~PC) +
  labs(y = "PCs' Value",
       x="True Quality")
```

From this figure, we can see that it's difficult to distinguish
the qualities by PCA

In order to assure this statement, we decide to use the median values to find which PC have ability to distinguish the qualities

```{r message=FALSE, warning=FALSE, echo=FALSE}
wine_data_pca %>%
  group_by(quality, PC) %>%
  summarise(median_feature = median(value)) %>%
  ggplot() +
  geom_point(aes(x=factor(quality), y=median_feature)) +
  facet_wrap(~PC, ncol=3)+
  labs(y = "Median Value of Principal Components",
       x="True Quality")
```
From this figure, we can see that PC2 and PC3 may help us to distinguish the qualities

```{r message=FALSE, warning=FALSE, echo=FALSE}
ggplot(wine_new) +
  geom_point(aes(x=PC2, y=PC3, color=factor(quality))) +
  labs(y = "PC3", x="PC2",
       color='True Quality')
```
However, according to this figure, we still can't use PCA to distinguish the quality successfully.

## Conclusion
Whether PCA or Clustering, they can distinguish the colors of wines successfully, but both of them can't distinguish the qualities of wines
Although PCA is better than Clustering, but it still doesn't have good ability to distinguish the qualities of wines
